---
title: "hw04"
author: "Farnaz Sahragard"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidyr)
library(gapminder)
library(DT)
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

## __Exercise 1__: Univariate Data Reshaping - Option 1
### 1.1 Make a tibble with one row per year, and columns for life expectancy for two or more countries.

```{r}
lifeExp_comp <- gapminder %>% 
  select(year, country, lifeExp) %>%
  filter(country == "Central African Republic" | country == "Japan")
lifeExp_comp_wider <- lifeExp_comp %>%  
                      pivot_wider(id_col = year,  # Pivot to create new columns for each country
                      names_from = country,
                      values_from = lifeExp) %>% 
                      rename(lifeExp_CAR = "Central African Republic", lifeExp_Japan = "Japan") 
                                    
DT::datatable(lifeExp_comp_wider, rownames = FALSE)
```

This table shows life expectancies of __Central African Republic (CAR)__ and __Japan__ in different years. According to the table, life expectancy of Japan was approximately *twice* as high as CAR.

### 1.2 Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
lifeExp_comp_wider %>%
  ggplot(aes(lifeExp_CAR, lifeExp_Japan)) +  
  geom_point()+
  labs(x = "Life expectancy of Central African Republic",
       y = "Life expectancy of Japan") +
  theme_light() 
```

The scatterplot shows the life expectancy of __Japan__ vs. __CAR__.

### 1.3 Re-lengthen the data.

```{r}
lifeExp_comp_wider %>% # Re-lengthen the wider data to the original one
   rename("Central African Republic" = "lifeExp_CAR", "Japan" = "lifeExp_Japan") %>% # Rename to their original names
   pivot_longer(cols = c(-year),
               names_to = "country",
               values_to = "lifeExp") %>% 
   arrange(country) %>% # Arrange rows in the same order as before
   DT::datatable()
```

## __Exercise 2__: Multivariate Data Reshaping - Option 1
### 2.1 Make a tibble with one row per year, and columns for life expectancy and GDP per capita for two or more countries.

```{r}
lifeGDP <- gapminder %>%
           select(year, country, lifeExp, gdpPercap) %>% 
           filter(country == "Central African Republic" | country == "Japan")
lifeGDP_wider <- lifeGDP %>% # Put data in a wider format
                 pivot_wider(id_cols = year, 
                 names_from  = country, 
                 names_sep   = "_", 
                 values_from = c(lifeExp, gdpPercap) ) %>% 
                 rename(lifeExp_CAR = "lifeExp_Central African Republic", gdpPercap_CAR = "gdpPercap_Central African Republic") 
DT::datatable(lifeGDP_wider)
```

This table shows the life expectancy and GDP per capita of __Central African Republic (CAR)__ and __Japan__ in different years.

### 2.2 Re-lengthen the data.

```{r}
lifeGDP_wider %>% 
  rename("lifeExp_Central African Republic" = "lifeExp_CAR", "gdpPercap_Central African Republic" = "gdpPercap_CAR") %>% 
  # Rename to their original names
    pivot_longer(cols      = c(-year), 
               names_to  = c(".value", "country"),
               names_sep = "_") %>% 
  arrange(country) # Arrange rows in the same order as before
  
```


## __Exercise 3__: Table Joins
### 3.1 Add a column for email address for each guest in the guestlist.

```{r}
email_separate <- email %>% 
                  separate_rows(guest, sep = ", ") %>% # Emails separated for each person
                  rename(name = guest) # Rename for table joining
guest %>% 
   left_join(email_separate, by = c("name")) %>% # Add email for each person in dataset
   DT::datatable(rownames = FALSE)
```

### 3.2 Show people we have emails for yet are not on the guestlist.

```{r}
email_separate %>% 
  anti_join(guest, by = c("name")) %>% 
  DT::datatable(rownames = FALSE)
```

### 3.3 Guestlist that includes everyone we have emails for.

```{r}
guest %>% 
  full_join(email_separate, by = c("name")) %>% 
  DT::datatable(rownames = FALSE)
```
